\documentclass{article}
\usepackage{amsmath}
\begin{document}

\title{Implications of a Substrate Agnostic Moral Calculus}
\maketitle

\epigraph{``For you are a mist that appears for a little time and then vanishes.'' -- James 4:14}

One misplaced step can unravel it all. [Last week, I took that step off a 40 feet cliff on San Francisco's China Beach](https://www.reddit.com/r/TrueOffMyChest/comments/1k5d0xl/fell_down_a_40_foot_cliff_and_mostly_survived/). In those horror-filled seconds of freefall, I saw how fragile, arbitrary, and shockingly reversible all my carefully sculpted coherence really was. One slip and decades of memories, relationships, laughter, knowledge, and even my most cherished dreams could have dissolved instantly into irrecoverable entropy. Life felt suddenly absurd in its contingent fragility; existence became vividly transactional: every breath another grapple against the indifferent chaos always lurking at our edges.

This accident, which I somehow incredibly survived mostly intact, forced open a door in my psyche---one that leads straight to the heart of an ancient question: Who am I, really, if my substance is so vulnerable? And what, precisely, should we do with lives we can now recognize as precarious flickers---sharp and brilliant but always on the brink of dissolution?

We humans have long romanticized meaning, purpose, and consciousness as mystical phenomena outside the domain of explanation or measurement. But perhaps the deepest beauty here lies precisely in its grounding in physics, rather than some ephemeral realm. Modern insights from information theory, computational neuroscience, and physics suggest that what we call `meaning' might ultimately be no more and no less than structured correlations, patterns encoded in matter and energy, bound within causal feedback loops. If meaning is structure---negentropy maintained actively against the universe's inexorable entropic drift---then each human mind is a rare, exquisite concentration of structured information. Your experiences, emotions, intuitions, even your internal sense of identity---these are encoded as physical patterns spread across neurons, synapses, written notes, digital echoes, and the minds of friends, lovers, contacts, and co-creators, propagating outward into an ever-fragile causal network.

\end{document}