\section{Formalizing Meaning: System, Agent, and Observer}

To formalize this intuition, consider a defined \textbf{System} ($S$), whose state we represent as $\mathbf{X}(t)$, changing over time. An \textbf{Agent} ($A$), perhaps a person like you or me, acts upon and within this system. We might also specify an \textbf{Observer} ($O$), whose viewpoint sets the frame for measurement (typically we assume an ``ideal'' objective observer). Within this framework, anything that exists and persists as structured correlations---memories, knowledge, neural patterns, digital archives---is at constant risk of erosion under the universal entropic drive toward randomness.

To measure this quantitatively, we define a measure of correlation or structured information $\mathcal{C}(\mathbf{X}(t))$. Several candidates exist, each offering a slightly different formal flavor but reflecting the underlying idea clearly:

\begin{itemize}
    \item \textbf{Negentropy} ($\mathcal{J}$), the deviation from maximum entropy:
    \begin{equation*}
        \mathcal{J} = H_{max} - H(\mathbf{X}(t))
    \end{equation*}
    A larger $\mathcal{J}$ indicates greater structure or reduced uncertainty about the system’s state.
    \item \textbf{Total Correlation (Multi-information $TC$)}, measuring redundancy or correlation across system components ($X_i$):
    \begin{equation*}
        TC(\mathbf{X}(t)) = \sum_i H(X_i(t)) - H(\mathbf{X}(t))
    \end{equation*}
    Greater total correlation means the system is more internally structured and interconnected.
\end{itemize}

Now, how does structured correlation change over time? Naturally, left alone, systems drift back towards chaos---their correlations degrade, bit by bit, through thermal fluctuations, random perturbations, or decay processes. But critically, agents like humans actively counteract these erosions. Formally, the change in structured correlation includes two competing terms: natural decay versus agent-driven structuring:

\begin{equation*}
\frac{d\mathcal{C}}{dt} = \frac{d\mathcal{C}}{dt}\Big|_{\text{natural}} + \frac{d\mathcal{C}}{dt}\Big|_{\text{agent}}
\end{equation*}

Typically, $\frac{d\mathcal{C}}{dt}\Big\vert_{\text{natural}} \leq 0$. Structure spontaneously breaks down; meaning vanishes in isolation. Agent-driven structuring---the intentional building or preservation of correlations over time---is thus precisely the rational measure we seek for ``generated meaning.''

We can now crisply formulate meaning generation rates ($\mathcal{M}_{\text{rate}}$) and accumulated meaning ($\mathcal{M}_{\text{total}}$) as follows:

\textbf{Definition 1: Rate of Meaning Generation ($\mathcal{M}_{\text{rate}}$)}

At any instant, the rate at which agent $A$ generates meaning within system $S$, relative to observer $O$, is:

\begin{equation*}
\mathcal{M}_{\text{rate}}(A, S, O, t) = \frac{d\mathcal{C}(\mathbf{X}_O(t))}{dt}\Big|_{\text{agent}} \quad (\text{bits/time})
\end{equation*}

Intuitively, this measures how quickly an agent adds structured information---meaning---into the world at a given moment.

\textbf{Definition 2: Accumulated Meaning ($\mathcal{M}_{\text{total}}$)}

Over any interval of lifetime activity $[t_0, t_f]$, an agent accumulates total meaning generated:

\begin{equation*}
\mathcal{M}_{\text{total}}(A, S, O, [t_0, t_f]) = \int_{t_0}^{t_f} \mathcal{M}_{\text{rate}}(A, S, O, t) \, dt \quad (\text{bits})
\end{equation*}

This integral corresponds directly to the total structured information the agent successfully implants into the world's fabric across its entire existence, standing defiantly against chaotic dissolution.

Understood through this formal lens, your existence---your most cherished memories, your imperfect relationships, your private longings, and your half-written theories---ceases to feel helplessly ephemeral or mysteriously ineffable. Instead, you become something profoundly concrete: a potent and quantifiable source of structured correlations, propagating coherence outward into digital archives, dialogues, institutions, or cultures. Your life trajectory can thus be formally measured as a cumulative informational legacy: the integral of your active preservation and amplification of ordered patterns against persistent entropy.
